---
title: "assignment_6"
author: "Nicole Szeluga"
date: "11/1/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Assignment 6: Relational data plus revisiting data wrangling and
visualization
================

## Instructions: Please read through this before you begin

  - This assignment is due by **10pm on Monday 11/02/20**.

  - Please **reproduce this markdown template**. Pay attention to all
    the formating in this file, including bullet points, bolded
    characters, inserted code chunks, headings, text colors, blank
    lines, and etc.

  - For exercise 1, use the appropriate `join` function to **join** some
    relational datasets, and use the resulting datasets to answer some
    questions.

  - For the second exercise in this assignment, you will use the data
    wrangling and visualization skills that you have learned in this
    class to reproduce a figure exactly as it is shown.

  - Have all your code embedded within the R markdown file, and show
    both of your **code** and **plots** in the knitted markdown file.

  - Use R Markdown functionalities to **hide messages and warnings when
    needed**. (Suggestion: messages and warnings can often be
    informative and important, so please examine them carefully and only
    turn them off when you finish the exercise).

  - Please name your R markdown file `assignment_6.Rmd` and the knitted
    markdown file `assignment_6.md`. Please upload both files using your
    personal GitHub repository for this class.

  - First, load all the required packages with the following code.
    Install them if they are not installed yet.

<!-- end list -->

```{r}
library(tidyverse)
library(knitr)
```


<br>

## Excercise 1: Join functions

**Acknowledgements:** The datasets and some of the questions for this
exercise has been borrowed (with permission) from Iain Carmichael’s
course [STOR 390: Introduction to Data
Science](https://idc9.github.io/stor390/#course_material).

<br>

#### Read in some relational datasets from GitHub


```{r}
toy1 <- read_csv('https://raw.githubusercontent.com/idc9/stor390/master/data/joins/toy_one.csv')
toy2 <- read_csv('https://raw.githubusercontent.com/idc9/stor390/master/data/joins/toy_two.csv')

## for 4.2
play1 <- read_csv('https://raw.githubusercontent.com/idc9/stor390/master/data/joins/play_one.csv')
play2 <- read_csv('https://raw.githubusercontent.com/idc9/stor390/master/data/joins/play_two.csv')

## for 4.3
banking_account_types <- read_csv('https://raw.githubusercontent.com/idc9/stor390/master/data/joins/banking_account_types.csv')
banking_transactions <- read_csv('https://raw.githubusercontent.com/idc9/stor390/master/data/joins/banking_transactions.csv')

## for 4.4
crime_arrests <- read_csv('https://raw.githubusercontent.com/idc9/stor390/master/data/joins/crime_arrests.csv')
crime_wealth <- read_csv('https://raw.githubusercontent.com/idc9/stor390/master/data/joins/crime_wealth.csv')

## for 4.5
titanic_outcomes <- read_csv('https://raw.githubusercontent.com/idc9/stor390/master/data/joins/titanic_outcomes.csv')
titanic_person_features <- read_csv('https://raw.githubusercontent.com/idc9/stor390/master/data/joins/titanic_person_features.csv')
```



<br>

#### 1.1 Join the `toy` datasets. Only keep rows that have matches in the other dataset. Show the resulting dataset using `kable()`.

```{r}
 rename(toy2,key = label) %>% 
  left_join(toy1, toy2, by = "key") %>% 
  na.omit() %>% 
  kable()
```


<br>

#### 1.2 Join the `play` datasets by `key` and `label`. Keep all rows from both datasets. Show the resulting dataset using `kable()`.

```{r}
full_join(play1, play2, by = c("key", "label")) %>% 
  kable()
```


<br>

#### 1.3 Plot the distribution of total transaction amount per account ID grouped by different account types with a boxplot shown below, using the `banking` datasets.

    ## `summarise()` regrouping output by 'id' (override with `.groups` argument)


```{r}
 
rename(banking_transactions, account_id = id) %>% 
  left_join(banking_account_types, banking_transactions, by = "account_id") %>% 
  group_by(account_id, account_type) %>% 
  summarise("total_transaction_amount" = sum(transaction_amount)) %>% 
  ggplot(mapping = aes( x = account_type, y = total_transaction_amount)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(height=0)
```

<br>

#### 1.4 Which state is in the `crime_wealth` dataset but not the `crime_arrests` dataset? Use a join function to answer this question.

```{r}
left_join( crime_wealth, crime_arrests, by = "State") %>% 
  kable()
```

The answer is the District of Columbia. 

<br>

## Exercise 2: Fuel Efficiency of Car Models

This exercise uses the `mtcars` dataset. It was extracted from the 1974
Motor Trend US magazine, and comprises fuel consumption and 10 aspects
of automobile design and performance for 32 automobiles (1973–74
models).

```{r}
kable(head(mtcars))
```

<br>

#### 2.1 Reproduce the following plot, which shows the miles per gallon (`mpg`) of car models on the x axis (see hints below). Different models are ordered on the y axis according to their `mpg` and their names are shown next to the data points. Also, the size of each data point maps to its horse power (`hp`), and the color maps to number of cylinders (`cyl`).

```{r}

mtcars %>% 
  rownames_to_column() %>% 
  mutate("mpg_order" = row_number(mpg)) %>% 
  ggplot(mapping = aes(x = mpg, y= mpg_order, color = factor(cyl), size = hp)) +
  geom_point() +
  geom_text(aes(label = rowname, hjust = -.15), size = 3.5, color = "black") +
  xlab("Miles per gallon fuel consumption") +
  ylab("") +
  labs(size = "Horsepower", color = "Number of cylinders") +
  theme(axis.line.y = element_blank(), panel.background = element_blank(), axis.line.x = element_line(), axis.ticks.y.left = element_blank(), axis.text.y = element_blank(), legend.key = element_blank()) +
  xlim(0,40) 
  
  

 
```


<br>

